Managing variability in software systems

### 2021-04-29 Paper.tex
work on feature modeling

\section{Variability modeling (1 page) JR}
As mentioned there are two main approaches for managing variability. For each of which we will discuss the advantages and disantages and create a broad overview over the different modeling approaches.\\\\
\textbf{Feature modeling}\\
Features in the context of feature modeling are behaviours, characetristics or functionalities which are observable by the user of the software system. They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}. A broadly used approach of displaying features and their relations are hierarchically structured diagrams which are called features diagrams.\\\\
A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders. This can be achieved by providing a overview of all functionalities which are offered to the user by the system. Due to the nature of the model regarding structure and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system. 
\\\\
\textbf{Decision modeling}\\
asdfasdf

@@

\begin{thebibliography}{00}
\bibitem{b1} S. Apel, D. Batory, C. Kästner and G. Saake, "Software Product Lines. In: Feature-Oriented Software Product Lines", Berlin Heidelberg, Springer (2013)
\bibitem{b2} J. van Gurp, J. Bosch and M. Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture", Amsterdam, Netherlands, 2001, pp. 45-54
\bibitem{b3} F. Bachmann, L. Bass, "Managing Variability in Software Architectures", Carnegie Mellon University (2001)
\bibitem{b4} M. Galster, D. Weyns, D. Tofan, M. Bartosz, P. Avgeriou, "Variability in Software Systems—A Systematic Literature Review", IEEE Transactions on Software Engineering (2014)
\bibitem{b5} T. Berger, J.-P. Steghöfer, T. Ziadi, J. Robin, and J. Martinez, "The state of adoption and the challenges of systematic variability management in industry", Empirical Software Engineering, vol. 25, pp. 1755-1797, 2020
\bibitem{b6}
K. Czarnecki, P. Grünbacher, R. Rabiser, K. Schmid., and A.
Wasowski, “Cool features and tough decisions: a comparison of
variability modeling approaches,” Proc. of the \textit{Sixth international
workshop on variability modeling of software-intensive systems
(VaMoS '12)}, ACM, Jan 2012, pp. 173-182.
\bibitem{b7}
Software Productivity Consortium Services Corporation,
Technical Report SPC-92019-CMC. \textit{Reuse-Driven Software
Processes Guidebook}, Version 02.00.03, 1993.
\bibitem{b8} K. C. Kang, S. Cohen, J. Hess, W. Novak, and S. Peterson,	 "Feature oriented domain analysis (FODA) feasibility study", Technical Report CMU/SEI-90-TR-21, Software Engineering Institute, Carnegie Mellon University, Pittburgh, PA, USA, 1990

\end{thebibliography}

>>>>

\section{Variability modeling (1 page) JR}
As mentioned there are two main approaches for managing variability. For each of which we will discuss the advantages and disadvantages and create a broad overview over the different modeling approaches.\\\\
\textbf{Feature modeling}\\
Features in the context of feature modeling are behaviours, characteristics or functionalities which are observable by the user of the software system. They can be used to define commonalities and differences between systems and their variants, as well as determine mandatory, optional or alternative aspects of them\cite{b8}. A broadly used approach of displaying features and their relations is hierarchically structured diagram which is called a features diagrams.\\\\
A big advantage of feature models, especially in the shape of feature diagrams, is a possible communication enhancement between stakeholders. This can be achieved by providing a overview of all functionalities which are offered to the user by the system. Due to the nature of the model regarding structure and due to the fact that it displays if and which planned features are already provided by another system or system family, this may improve the cost and time estimation for developing the software system. \\\\
The essence of feature modeling is the splitting of features into distinct objects which have a certain relation to each other. There are several types of features mentioned in \cite{b8}, named \textit{mandatory, optional and alternative features}. Generally speaking common attributes of systems are described by mandatory features, while alternative or optional features define the variability in feature models.\cite{b9}\\\\ 
If a feature is mandatory it has to be selected if its parent feature is selected, the type of the parent feature does not matter. A optional feature can be selected freely but there may be constraints from composition rules (e.g. a feature can only be choosen as part of a feature sets). Alternative features are can be pictured as a choice box, so one can choose \textit{one} item from a feature set.\\\\
This is the approach which is used in the original FODA paper, which was extended with the introduction of cardinalities for alternative features. \cite{b10} This expansion grants the ability to create minimum and maximum constraints for choosing alternative features from a feature set, which helps with creating complex models.

\textbf{Decision modeling}\\
asdfasdf

@@

\begin{thebibliography}{00}
\bibitem{b1} S. Apel, D. Batory, C. Kästner and G. Saake, "Software Product Lines. In: Feature-Oriented Software Product Lines", Berlin Heidelberg, Springer (2013)
\bibitem{b2} J. van Gurp, J. Bosch and M. Svahnberg, "On the notion of variability in software product lines," Proceedings Working IEEE/IFIP Conference on Software Architecture", Amsterdam, Netherlands, 2001, pp. 45-54
\bibitem{b3} F. Bachmann, L. Bass, "Managing Variability in Software Architectures", Carnegie Mellon University (2001)
\bibitem{b4} M. Galster, D. Weyns, D. Tofan, M. Bartosz, P. Avgeriou, "Variability in Software Systems—A Systematic Literature Review", IEEE Transactions on Software Engineering (2014)
\bibitem{b5} T. Berger, J.-P. Steghöfer, T. Ziadi, J. Robin, and J. Martinez, "The state of adoption and the challenges of systematic variability management in industry", Empirical Software Engineering, vol. 25, pp. 1755-1797, 2020
\bibitem{b6}
K. Czarnecki, P. Grünbacher, R. Rabiser, K. Schmid., and A.
Wasowski, “Cool features and tough decisions: a comparison of
variability modeling approaches,” Proc. of the \textit{Sixth international
workshop on variability modeling of software-intensive systems
(VaMoS '12)}, ACM, Jan 2012, pp. 173-182.
\bibitem{b7}
Software Productivity Consortium Services Corporation,
Technical Report SPC-92019-CMC. \textit{Reuse-Driven Software
Processes Guidebook}, Version 02.00.03, 1993.
\bibitem{b8} K. C. Kang, S. Cohen, J. Hess, W. Novak, and S. Peterson,	 "Feature oriented domain analysis (FODA) feasibility study", Technical Report CMU/SEI-90-TR-21, Software Engineering Institute, Carnegie Mellon University, Pittburgh, PA, USA, 1990
\bibitem{b9} A. Metzger, and K. Pohl, "Software product line engineering and variability management: Achievements and challenges," \textit{Proc. of the Future of Software Engineering (FOSE 2014)}, ACM, 2014, pp. 70-84, doi:10.1145/2593882.2593888.
\bibitem{b10} M. Riebisch, K. Böllert, D. Streitferdt, and I. Philippow, "Extending Feature Diagrams with UML Multiplicities", \textit{6th World Conference on Integrated Design \& Process Technology (IDPT2002)}, vol. 23, 2002, pp. 1-7.
\end{thebibliography}
